---
title: ""
author: ""
date: ""
params:
  asset1: NULL
  asset2: NULL
  entry: NULL
  exit: NULL
  crypto_data: NULL
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: null
    highlight: zenburn
    css: custom_report_style.css
    self_contained: true
    df_print: paged
---

```{r topbar_html, echo=FALSE, results='asis'}
htmltools::HTML('
<div class="topbar">
  <span class="topbar-title">Reports:</span>
  <div class="topbar-links">
    <a href="BTCUSD_DOGEUSD_report.html">BTC/USD - DOGE/USD</a>
    <a href="BTCUSD_XRPUSD_report.html">BTC/USD - XRP/USD</a>
    <a href="DOGEUSD_ADAUSD_report.html">DOGE/USD - ADA/USD</a>
    <a href="DOGEUSD_XRPUSD_report.html">DOGE/USD - XRP/USD</a>
    <a href="ETHUSD_ADAUSD_report.html">ETH/USD - ADA/USD</a>
    <a href="ETHUSD_DOGEUSD_report.html">ETH/USD - DOGE/USD</a>
    <a href="ETHUSD_SOLUSD_report.html">ETH/USD - SOL/USD</a>
    <a href="LTCUSD_ADAUSD_report.html">LTC/USD - ADA/USD</a>
    <a href="SOLUSD_DOGEUSD_report.html">SOL/USD - DOGE/USD</a>
    <a href="SOLUSD_XRPUSD_report.html">SOL/USD - XRP/USD</a>
    <a href="XRPUSD_ADAUSD_report.html">XRP/USD - ADA/USD</a>
    <a href="XRPUSD_LTCUSD_report.html">XRP/USD - LTC/USD</a>
  </div>
</div>
')

```

<!-- Inizio contenuto principale -->
<div class="container">

<h1 style="margin-bottom: 0.5rem;">Pairs Trading Report: `r params$asset1` / `r params$asset2`</h1>
<p style="margin-top: 0; margin-bottom: 0;"><strong>Nicolò Blasco Pecoraro</strong></p>
<p style="margin-top: 0;"><strong>`r format(Sys.Date(), "%d %B %Y")`</strong></p>

```{r setup, include=FALSE}
library(PerformanceAnalytics)
library(xts)
library(ggplot2)
library(scales)
library(knitr)
library(htmltools)

options(digits = 3, scipen = 999)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```
## 1. Strategy Overview

This report analyzes the performance of a pairs trading strategy between the following assets:

- **Asset 1**: `r params$asset1`  
- **Asset 2**: `r params$asset2`

We use daily log-prices and model their spread using a linear regression.  
Entry and exit thresholds are based on z-score deviations:

- **Entry Threshold**: `r params$entry`  
- **Exit Threshold**: `r params$exit`


## 2. Strategy Logic & Execution

This section outlines the core logic behind the pairs trading strategy.

We estimate the relationship between the two assets using a linear regression on log-prices.  
The spread (i.e., residual) from this model is then used to compute a **z-score**,  
which serves as the signal for both entry and exit decisions.

Based on this logic, we apply a simple rule-based trading strategy:

- **Enter Long**: when the z-score falls below the negative entry threshold.
- **Enter Short**: when the z-score rises above the positive entry threshold.
- **Exit**: when the z-score reverts within the exit threshold band (i.e., close to 0).

The trading positions and their resulting returns are calculated accordingly.


```{r strategy_logic}
# Data from Crypto
x <- params$crypto_data[[params$asset1]]
y <- params$crypto_data[[params$asset2]]

# Data Merge and NA omission
df <- na.omit(merge(x, y))

# Log Calculation
log_x <- log(df[, 1])
log_y <- log(df[, 2])

# Linear Regression
model <- lm(log_x ~ log_y)
spread <- residuals(model)

# Z-score
spread_ret <- diff(log_x - coef(model)[2] * log_y)
zscore <- (spread - mean(spread)) / sd(spread)

# Trading Strategy
positions <- rep(0, length(zscore))
for (j in 2:length(zscore)) {
  if (zscore[j - 1] > params$entry) {
    positions[j] <- -1
  } else if (zscore[j - 1] < -params$entry) {
    positions[j] <- 1
  } else if (abs(zscore[j - 1]) < params$exit) {
    positions[j] <- 0
  } else {
    positions[j] <- positions[j - 1]
  }
}
``` 

```{r performance_returns}
positions_xts <- xts(positions, order.by = index(zscore))
strategy_ret <- lag(positions_xts) * spread_ret
strategy_ret <- na.omit(strategy_ret)
```


## 3. Strategy Performance


```{r performance_summary, fig.align='left', fig.width=8, fig.height=5}
charts.PerformanceSummary(
  strategy_ret,
  main = paste("Performance Summary:", params$asset1, "/", params$asset2),
  colorset = c("black", "darkgray", "firebrick")
)
```

### Performance Metrics Table


```{r performance_metrics}
total_ret <- round((prod(1 + strategy_ret) - 1) * 100, 2)
sharpe <- round(SharpeRatio.annualized(strategy_ret)[[1]], 2)
drawdown <- round(maxDrawdown(strategy_ret), 4)

perf_table <- setNames(data.frame(
  total_ret,
  sharpe,
  drawdown
), c("Total Return (%)", "Annualized Sharpe", "Max Drawdown"))


knitr::kable(perf_table, caption = "Strategy Performance Metrics")
```
## 4. Interpretation

This section provides a qualitative interpretation of the strategy’s performance, using the Sharpe Ratio as a key metric.


```{r interpretation_comment, results='asis', echo=FALSE}
if (sharpe >= 1) {
  cat("<strong>Excellent Risk-Adjusted Performance</strong><br><br>")
  cat("The strategy demonstrates a strong capacity to generate excess returns relative to its risk. A Sharpe Ratio above 1 typically suggests that the strategy is both consistent and profitable, and may be considered for real-world implementation. It indicates a high level of confidence in the statistical edge of the model.\n")
} else if (sharpe >= 0.5) {
  cat("\\textbf{Moderate Risk-Adjusted Returns}\n\n")
  cat("The strategy offers reasonable potential for generating returns that compensate for the risk taken. While the Sharpe Ratio is below the ideal threshold of 1, the model could still be viable under certain market conditions, especially if further optimized or paired with complementary strategies to reduce drawdown.\n")
} else {
  cat("\\textbf{Weak Risk-Adjusted Performance}\n\n")
  cat("The strategy currently exhibits low performance relative to its volatility. A Sharpe Ratio below 0.5 suggests that risks are not being adequately compensated by returns. Caution is advised before deploying this strategy in practice—further tuning or alternative logic may be required.\n")
}
```


## 5. Volatility Profile & Risk Evaluation

This section evaluates the volatility structure of the strategy returns. Here are reported key indicators such as standard deviation, realized volatility, annualized volatility, mean return per trade, and number of trades.  
These indicators help assess the overall risk profile of the pair.

```{r volatility_metrics, fig.align='left', fig.width=18, fig.height=6}
# Compute volatility indicators
sd_ret <- if (length(strategy_ret) > 1) sd(strategy_ret) else NA
realized_vol <- if (length(strategy_ret) > 1) sqrt(sum(strategy_ret^2)) else NA
annual_vol <- if (!is.na(sd_ret)) sd_ret * sqrt(252) else NA
mean_ret <- if (length(strategy_ret) > 1) mean(strategy_ret) else NA
num_trades <- if (length(positions_xts) > 1) sum(diff(positions_xts) != 0, na.rm = TRUE) else 0

vol_table <- data.frame(
  "Standard Deviation" = round(sd_ret, 5),
  "Realized Volatility" = round(realized_vol, 5),
  "Annualized Volatility" = round(annual_vol, 5),
  "Mean Return per Trade" = round(mean_ret, 5),
  "Number of Trades" = num_trades
)

knitr::kable(vol_table, caption = "Volatility Metrics of the Strategy")
```

```{r}
comment <- if (is.na(annual_vol)) {
  "The strategy did not generate enough return observations to evaluate volatility. This may indicate that no signals were triggered or the spread remained highly stable throughout the period."
} else if (annual_vol < 0.4) {
  "The strategy shows **very low volatility**, indicating a highly stable spread and conservative signal behavior. This is ideal for low-risk trading environments."
} else if (annual_vol < 0.7) {
  "The strategy exhibits **moderate volatility**, suggesting a balanced exposure profile. This level of risk is generally acceptable in practical applications."
} else if (annual_vol < 1) {
  "The strategy has **elevated volatility**, which may be acceptable if supported by sufficiently high returns and consistent signal reliability."
} else {
  "The strategy demonstrates **high or excessive volatility**, raising concerns about robustness. Such strategies may require tighter stop-losses or additional risk filters to be viable."
}

knitr::asis_output(paste("### Volatility Commentary\n\n", comment))

```

```{r}
z_df <- data.frame(Date = index(zscore), Z = coredata(zscore))
z_df$Signal <- factor(positions, levels = c(-1, 0, 1), labels = c("Short", "Flat", "Long"))

ggplot(z_df, aes(x = Date, y = Z)) +
  geom_line(color = "steelblue") +
  geom_hline(yintercept = c(-params$entry, params$entry), linetype = "dashed", color = "red") +
  geom_hline(yintercept = c(-params$exit, params$exit), linetype = "dashed", color = "orange") +
  geom_point(data = subset(z_df, Signal == "Long"), aes(y = Z), color = "limegreen", size = 2) +
  geom_point(data = subset(z_df, Signal == "Short"), aes(y = Z), color = "orchid", size = 2) +
  labs(
    title = paste("Z-Score & Trading Signals:", params$asset1, "/", params$asset2),
    subtitle = paste("Entry:", params$entry, "| Exit:", params$exit),
    y = "Z-Score",
    x = "Date"
  ) +
  theme_minimal()

```


## 6. Conclusion

This report demonstrated the use of a **mean-reversion pairs trading strategy** on selected cryptocurrency pairs.

We applied a **dynamic entry-exit logic** based on z-scores, measured key risk-adjusted metrics, and generated performance summaries.

This framework is designed to be **reusable**.  
By uploading a different crypto pair and running the optimization pipeline, a new report can be produced automatically.

## Usefull Links

<div class="footer-links"> <a href="https://finance.yahoo.com/cryptocurrencies/" target="_blank"> Yahoo Finance Crypto</a> | <a href="https://coinmarketcap.com/" target="_blank"> CoinMarketCap</a> | <a href="https://www.tradingview.com/markets/cryptocurrencies/prices-all/" target="_blank"> TradingView</a> </div> <!-- close .container -->``




