---
title: "Pairs Trading Report: `r params$asset1` / `r params$asset2`"
author: "Nicolò Blasco Pecoraro"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: null
    highlight: zenburn
    css: custom_report_style.css
params:
  asset1: NULL
  asset2: NULL
  entry: NULL
  exit: NULL
  crypto_data: NULL
---
<!-- Sidebar per navigazione report -->
<div class="sidebar">
  <h2>📊 Reports</h2>
  <ul>
    <li><a href="BTCUSD_DOGEUSD_report.html">BTC/USD - DOGE/USD</a></li>
    <li><a href="BTCUSD_XRPUSD_report.html">BTC/USD - XRP/USD</a></li>
    <li><a href="DOGEUSD_ADAUSD_report.html">DOGE/USD - ADA/USD</a></li>
    <li><a href="DOGEUSD_XRPUSD_report.html">DOGE/USD - XRP/USD</a></li>
    <li><a href="ETHUSD_ADAUSD_report.html">ETH/USD - ADA/USD</a></li>
    <li><a href="ETHUSD_DOGEUSD_report.html">ETH/USD - DOGE/USD</a></li>
    <li><a href="ETHUSD_SOLUSD_report.html">ETH/USD - SOL/USD</a></li>
    <li><a href="LTCUSD_ADAUSD_report.html">LTC/USD - ADA/USD</a></li>
    <li><a href="SOLUSD_DOGEUSD_report.html">SOL/USD - DOGE/USD</a></li>
    <li><a href="SOLUSD_XRPUSD_report.html">SOL/USD - XRP/USD</a></li>
    <li><a href="XRPUSD_ADAUSD_report.html">XRP/USD - ADA/USD</a></li>
    <li><a href="XRPUSD_LTCUSD_report.html">XRP/USD - LTC/USD</a></li>
  </ul>
</div>

<!-- Inizio contenuto principale -->
<div class="container">

```{r setup, include=FALSE}
library(PerformanceAnalytics)
library(xts)
library(ggplot2)
library(scales)
library(knitr)

options(digits = 3, scipen = 999)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
## Strategy Overview

You are analyzing the pair:  
**`r params$asset1` / `r params$asset2`**

- Entry Threshold: `r params$entry`  
- Exit Threshold: `r params$exit`

---

## Strategy Performance

```{r strategy}
x <- params$crypto_data[[params$asset1]]
y <- params$crypto_data[[params$asset2]]
df <- na.omit(merge(x, y))

log_x <- log(df[, 1])
log_y <- log(df[, 2])
model <- lm(log_x ~ log_y)
spread <- residuals(model)
spread_ret <- diff(log_x - coef(model)[2] * log_y)
zscore <- (spread - mean(spread)) / sd(spread)

positions <- rep(0, length(zscore))
for (j in 2:length(zscore)) {
  if (zscore[j - 1] > params$entry) {
    positions[j] <- -1
  } else if (zscore[j - 1] < -params$entry) {
    positions[j] <- 1
  } else if (abs(zscore[j - 1]) < params$exit) {
    positions[j] <- 0
  } else {
    positions[j] <- positions[j - 1]
  }
}

positions_xts <- xts(positions, order.by = index(zscore))
strategy_ret <- lag(positions_xts) * spread_ret
strategy_ret <- na.omit(strategy_ret)

charts.PerformanceSummary(strategy_ret,
  main = paste("Performance Summary:", params$asset1, "/", params$asset2))
